
# Math - Percentage of Whole Calculations 

Here are a few refresher exercises for doing math calculations within the code.


```{r}
library(tidyverse)
```

We'll import a dataset of median household income in Baltimore for the 2010, 2016 and 2020 surveys. The data is in the data folder in this chapter's pre-lab directory. We'll use this to explore ways to create new information from existing data.

/Users/robwells/Code/data_journalism_class/04_labs/lab_05/pre_lab_05/baltcity_income_clean.csv

```{r}

baltcity_income<- read_csv("https://raw.githubusercontent.com/wellsdata/data_journalism_class/main/04_labs/lab_05/pre_lab_05/baltcity_income_clean.csv") %>% 
  as.data.frame()


```
Let's see which neighborhoods are the largest on a percentage basis in 2020 

```{r}
baltcity_income_2020 <- baltcity_income %>%
  select(Census, Neighborhood, x2020) %>%
  mutate(pct_whole = scales::percent(x2020 / sum(x2020, na.rm = TRUE), accuracy = 0.2)) %>% 
  arrange(desc(pct_whole))

```

Create a Diff_Income column and then a percentage change column for 2020 vs 2010

```{r}
baltcity_income %>%
  select(Census, Neighborhood, x2010, x2020) %>%
    mutate(Diff_Income = (x2020-x2010)) %>% 
    mutate(Diff_Pct_2020 = (x2020-x2010)/x2010)
  
```
